---
title: "ANLY 699 Cluster Assignment"
author: "Gaurav Gade"
date: "10/24/2021"
output: pdf_document
---

```{r setup, include=FALSE}
# install.packages("xts")
# install.packages("dplyr")


library(dplyr)
library(xts)
library(cluster)
getwd()
setwd("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets")



```

```{r Data preparation, cleaning and normalization}
#========DATA PREP==========================

    
#Import data for DeFi Tokens and BTC - BTC, CARDANO, CHAINLINK, POLKADOT, SOLANA
    
#Import data for each of the four securities - BTC
    data.BTC <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/BTC-USD.csv",header=TRUE)
    date <- as.Date(data.BTC$Date,format="%Y-%m-%d")  
    data.BTC <- cbind(date, data.BTC[,-1])
    data.BTC <- data.BTC[order(data.BTC$date),]
    data.BTC <- xts(data.BTC[,2:7],order.by=data.BTC[,1])
    names(data.BTC) <- paste(c("BTC.Open","BTC.High","BTC.Low","BTC.Close","BTC.Adjusted","BTC.Volume"))
    data.BTC[c(1:3,nrow(data.BTC)),]

#Import data for each of the four securities - Cardano
    data.ADA <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/ADA-USD.csv",header=TRUE)
    date <- as.Date(data.ADA$Date,format="%Y-%m-%d")  
    data.ADA <- cbind(date, data.ADA[,-1])
    data.ADA <- data.ADA[order(data.ADA$date),]
    data.ADA <- xts(data.ADA[,2:7],order.by=data.ADA[,1])
    names(data.ADA) <- paste(c("ADA.Open","ADA.High","ADA.Low","ADA.Close","ADA.Adjusted","ADA.Volume"))
    data.ADA[c(1:3,nrow(data.ADA)),]

#Import data for each of the four securities - ChainLINK (LINK)
    data.LINK <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/LINK-USD.csv",header=TRUE)
    date <- as.Date(data.LINK$Date,format="%Y-%m-%d")  
    data.LINK <- cbind(date, data.LINK[,-1])
    data.LINK <- data.LINK[order(data.LINK$date),]
    data.LINK <- xts(data.LINK[,2:7],order.by=data.LINK[,1])
    names(data.LINK) <- paste(c("LINK.Open","LINK.High","LINK.Low","LINK.Close","LINK.Adjusted","LINK.Volume"))
    data.LINK[c(1:3,nrow(data.LINK)),]

#Import data for each of the four securities - POLKADOT (DOT)
    data.DOT <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/DOT1-USD.csv",header=TRUE)
    date <- as.Date(data.DOT$Date,format="%Y-%m-%d")  
    data.DOT <- cbind(date, data.DOT[,-1])
    data.DOT <- data.DOT[order(data.DOT$date),]
    data.DOT <- xts(data.DOT[,2:7],order.by=data.DOT[,1])
    names(data.DOT) <- paste(c("DOT.Open","DOT.High","DOT.Low","DOT.Close","DOT.Adjusted","DOT.Volume"))
    data.DOT[c(1:3,nrow(data.DOT)),]
    
#Import data for each of the four securities - SOLANA (SOL)
    data.SOL <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/SOL1-USD.csv",header=TRUE)
    date <- as.Date(data.SOL$Date,format="%Y-%m-%d")  
    data.SOL <- cbind(date, data.SOL[,-1])
    data.SOL <- data.SOL[order(data.SOL$date),]
    data.SOL <- xts(data.SOL[,2:7],order.by=data.SOL[,1])
    names(data.SOL) <- paste(c("SOL.Open","SOL.High","SOL.Low","SOL.Close","SOL.Adjusted","SOL.Volume"))
    data.SOL[c(1:3,nrow(data.SOL)),]
    
    
    data.ADA <- subset(data.ADA,
                        index(data.ADA)>"2021-03-31")
    
    data.BTC <- subset(data.BTC,
                        index(data.BTC)>"2021-03-31")
    
    data.DOT <- subset(data.DOT,
                        index(data.DOT)>"2021-03-31")
    
    data.LINK <- subset(data.LINK,
                        index(data.LINK)>"2021-03-31")
    
    data.SOL <- subset(data.SOL,
                        index(data.SOL)>"2021-03-31")
    
    
    
#Combine Data Into one Data Object
    Close.Prices <- cbind(data.BTC$BTC.Close, data.ADA$ADA.Close,data.LINK$LINK.Close,data.SOL$SOL.Close, data.DOT$DOT.Close)
    Close.Prices[c(1:3,nrow(Close.Prices)),]


    
#Convert Data into a data.frame
    
    multi.df <- cbind(index(Close.Prices),data.frame(Close.Prices))
    names(multi.df) <- paste(c("date","BTC","ADA","LINK","DOT","SOL"))
    rownames(multi.df) <- seq(1,nrow(multi.df),1)
    multi.df[c(1:3,nrow(multi.df)),]

    
    multi.df$ADA <- as.numeric(multi.df$ADA)
    multi.df$BTC <- as.numeric(multi.df$BTC)
    multi.df$LINK <- as.numeric(multi.df$LINK)
    multi.df$DOT <- as.numeric(multi.df$DOT)
    multi.df$SOL <- as.numeric(multi.df$SOL)
  
    
    multi.df <- na.omit(multi.df)
    
    
    #Calculate Normalized Values for Each Security
#Create an index for each security with values that equal the price of the 	security on each day divided by the security's price on December 31, 	2010.

  
    
    
    multi.df$BTC.idx <- multi.df$BTC/multi.df$BTC[1]
    multi.df$ADA.idx <- multi.df$ADA/multi.df$ADA[1]
    multi.df$LINK.idx <- multi.df$LINK/multi.df$LINK[1]
    multi.df$SOL.idx <- multi.df$SOL/multi.df$SOL[1]
    multi.df$DOT.idx <- multi.df$DOT/multi.df$DOT[1]
    
    options(digits=5)   # this option remains in place until end of session
    multi.df[c(1:3,nrow(multi.df)),]    



```

```{r Capital Appreciation, echo=FALSE}
#################################################    
    
#Plot the Capital Appreciation of Each Security
#For the x variable use date, for y use Qualcomm Index

    
    multi <- data.ADA[,5]
    head(multi)
    multi <- merge(multi,data.BTC[,5])
    multi <- merge(multi,data.LINK[,5])
    multi <- merge(multi,data.SOL[,5])
    multi <- merge(multi,data.DOT[,5])
    
    
    multi[c(1:3,nrow(multi)),]
#============PLOTS======================
                
                    plot(x=multi.df$date,
                         y=multi.df$BTC.idx,
                         type="l",
                         xlab="Date",
                         ylab="Value of Investment ($)",
                         col="green",
                         lty=1,
                         lwd=1,
                         ylim = c(-2,6),
                         main="Value of $1 Investment in
                         ADA, BTC, LINK, SOL and DOT from 07-2010 - 07-2020")
                
                    #Adding the lines for the other 5 securities
                    
                       lines(x=multi.df$date,
                          y=multi.df$ADA.idx,
                          col="dark green",
                          lty=1,
                          lwd=1)
                    
                    
                    lines(x=multi.df$date,
                          y=multi.df$LINK.idx,
                          col="dark red",
                          lty=1,
                          lwd=1)
                    
                    lines(x=multi.df$date,
                          y=multi.df$DOT.idx,
                          col="dark blue",
                          lty=1,
                          lwd=1)
                    
                    lines(x=multi.df$date,
                          y=multi.df$SOL.idx,
                          col="blue",
                          lty=1,
                          lwd=1)
                    #Adding a line to denote the starting investment value of $1
                    
                    abline(h=1,lty=2,col="black")
                    
                    #Adding a legend
                    legend("topleft",
                           c("BTC","ADA","LINK","DOT","SOL"),
                           col=c("dark green","dark red","blue","dark blue"),
                           lty=c(1,1,1,1,1),
                           lwd=c(3,3,3,3,3))
                    
                  #The plot cuts off the values of some securities.
                    y.range <- range(multi.df[,6:9])  # find minimum/maximum of all 6 securities
                    y.range    
```

K-Means 50pts
  Find the optimal number of clusters (elbow, gap or silhouette methods). Plot the Figure. 10pts
  Explain how you decided on the optimal number. 10pts
  Perform the K-Means cluster analysis. Do not print results. 10pts
  Visualize cluster. Consider adding Y labels (see reading and practice for code). 10pts
  Provide your interpretation of the cluster visualization. Does it provide you with additional insights for your data? 10pts
  
```{r K-Means}

#K Means Clustering

#Set seed for reproducability
set.seed(123)
         
         df <- multi.df[,-c(1:6)]
        dft <- t(df)         
#na.omit(dft)

#measure.dist = dist(dft, method = "euclidean") 
#change method to maximum or manhattan 
#measure.dist

```
## Silhouette

- Using `sapply` calculate the average silhouette distances for 2 to n-1 clusters on only the second cluster analysis.

```{r silhouette}
##r chunk
measure.dist = dist(df, method = "euclidean")
measure.hc = hclust(measure.dist, method = "ward.D2")
sapply(2:5, #we can run 2 to n-1 clusters 
function(x) summary( 
silhouette(cutree(measure.hc, k = x), 
measure.dist)) #find the widths 
)

silhouette(cutree(measure.hc, k = 3), 
measure.dist)

#Determine optimal number of clusters
set.seed(1234)
if(!require('factoextra')){install.packages('factoextra')}
library(factoextra)
#Elbow method
fviz_nbclust(df, kmeans, method = "wss", nstart=20)


#Silhoutte method
fviz_nbclust(df, kmeans, method = "silhouette", nstart=20)



```
We can infer that the optimal number of clusters for the dataset = 2. 


Based on the cluster, we can interpret that Polkadot's stock performance varies from that of the other 4 crypto currencies. 

We observe that in 2021, BTC and SOlana are strongly correlated.

Chain Link is closer to the cluster of BTC and Sol.
Cardano has largely been moving in a direction that is slightly aligned to the cluster of LINK, BTC and SOL
but is not correlated significantly with Polkadot.

```{r K-Means2}

#Calculate with final number of clusters
set.seed(1234)
kMeansResult = kmeans(df, centers = 2)

#View results


#kMeansResult$cluster


```
#Results have intentionally not been printed. Based on the result, we can see that every observation has been assigned to a cluster 1 or 2. 

Hierarchical Cluster

Build hierarchical cluster. Explain what linkage is used (you can try several linkage methods).

Using two different linkage methods, I have built the hierarchical clusters using Ward method and Average linkage.
I notice similar dendograms being generated.

Plot hierarchical cluster.
How many clusters do you observe. I notice 2 main clusters. I used cuttree to improve the performance and
visual appeal.

Interpret the results. Does it provide you with additional insights? 

In the case of hierarchical cluster, we notice similar results to that of K-means. 



(Make sure to use a proper academic writing style to describe results - see the required book on academic writing)




```{r HClust}
#Set seed for reproducability
set.seed(123)
         
         df <- multi.df[,-c(1:6)]
                
#na.omit(dft)

measure.dist = dist(df, method = "euclidean")

measure.hc = hclust(measure.dist, method = "ward.D2") 
plot(measure.hc, hang = -1)

clusterCut <- cutree(measure.hc, 2)

#Using a difference Linkage method
clusters <- hclust(dist(df), method = 'average')
plot(clusters)

clusterCut <- cutree(clusters, 2)

```

```{r replot}
##r chunk
{
  plot(measure.hc, hang = -1) 
  rect.hclust(measure.hc, k = 2)
}


```