---
title: "Anly 699 Assignment 4"
author: "Gaurav Gade"
date: "10/18/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(ggplot2)
library(modelr)

```

```{r Data preparation, cleaning and normalization}
#========DATA PREP==========================

    
#Import data for DeFi Tokens and BTC - BTC, CARDANO, CHAINLINK, POLKADOT, SOLANA
    
#Import data for each of the four securities - BTC
    data.BTC <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/BTC-USD.csv",header=TRUE)
    date <- as.Date(data.BTC$Date,format="%Y-%m-%d")  
    data.BTC <- cbind(date, data.BTC[,-1])
    data.BTC <- data.BTC[order(data.BTC$date),]
    data.BTC <- xts(data.BTC[,2:7],order.by=data.BTC[,1])
    names(data.BTC) <- paste(c("BTC.Open","BTC.High","BTC.Low","BTC.Close","BTC.Adjusted","BTC.Volume"))
    data.BTC[c(1:3,nrow(data.BTC)),]

#Import data for each of the four securities - Cardano
    data.ADA <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/ADA-USD.csv",header=TRUE)
    date <- as.Date(data.ADA$Date,format="%Y-%m-%d")  
    data.ADA <- cbind(date, data.ADA[,-1])
    data.ADA <- data.ADA[order(data.ADA$date),]
    data.ADA <- xts(data.ADA[,2:7],order.by=data.ADA[,1])
    names(data.ADA) <- paste(c("ADA.Open","ADA.High","ADA.Low","ADA.Close","ADA.Adjusted","ADA.Volume"))
    data.ADA[c(1:3,nrow(data.ADA)),]

#Import data for each of the four securities - ChainLINK (LINK)
    data.LINK <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/LINK-USD.csv",header=TRUE)
    date <- as.Date(data.LINK$Date,format="%Y-%m-%d")  
    data.LINK <- cbind(date, data.LINK[,-1])
    data.LINK <- data.LINK[order(data.LINK$date),]
    data.LINK <- xts(data.LINK[,2:7],order.by=data.LINK[,1])
    names(data.LINK) <- paste(c("LINK.Open","LINK.High","LINK.Low","LINK.Close","LINK.Adjusted","LINK.Volume"))
    data.LINK[c(1:3,nrow(data.LINK)),]

#Import data for each of the four securities - POLKADOT (DOT)
    data.DOT <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/DOT1-USD.csv",header=TRUE)
    date <- as.Date(data.DOT$Date,format="%Y-%m-%d")  
    data.DOT <- cbind(date, data.DOT[,-1])
    data.DOT <- data.DOT[order(data.DOT$date),]
    data.DOT <- xts(data.DOT[,2:7],order.by=data.DOT[,1])
    names(data.DOT) <- paste(c("DOT.Open","DOT.High","DOT.Low","DOT.Close","DOT.Adjusted","DOT.Volume"))
    data.DOT[c(1:3,nrow(data.DOT)),]
    
#Import data for each of the four securities - SOLANA (SOL)
    data.SOL <- read.csv("C:/Users/nagar/Documents/ANLY 699 final project/Defi Datasets/SOL1-USD.csv",header=TRUE)
    date <- as.Date(data.SOL$Date,format="%Y-%m-%d")  
    data.SOL <- cbind(date, data.SOL[,-1])
    data.SOL <- data.SOL[order(data.SOL$date),]
    data.SOL <- xts(data.SOL[,2:7],order.by=data.SOL[,1])
    names(data.SOL) <- paste(c("SOL.Open","SOL.High","SOL.Low","SOL.Close","SOL.Adjusted","SOL.Volume"))
    data.SOL[c(1:3,nrow(data.SOL)),]
    
    
    data.ADA <- subset(data.ADA,
                        index(data.ADA)>"2021-03-31" & index(data.ADA) <"2021-10-23" )

    data.BTC <- subset(data.BTC,
                        index(data.BTC)>"2021-03-31"  & index(data.BTC) <"2021-10-23")

    data.DOT <- subset(data.DOT,
                        index(data.DOT)>"2021-03-31"  & index(data.DOT) <"2021-10-23")

    data.LINK <- subset(data.LINK,
                        index(data.LINK)>"2021-03-31"  & index(data.LINK) <"2021-10-23")

    data.SOL <- subset(data.SOL,
                        index(data.SOL)>"2021-03-31"  & index(data.SOL) <"2021-10-23")


    #summary(data.ADA)
```   

```{r Combine Data}    
    #Combine Data Into one Data Object
    Close.Prices <- cbind(data.BTC$BTC.Close, data.ADA$ADA.Close,data.LINK$LINK.Close,data.SOL$SOL.Close, data.DOT$DOT.Close)
    Close.Prices[c(1:3,nrow(Close.Prices)),]


    
#Convert Data into a data.frame
    
    multi.df <- cbind(index(Close.Prices),data.frame(Close.Prices))
    names(multi.df) <- paste(c("date","BTC","ADA","LINK","DOT","SOL"))
    rownames(multi.df) <- seq(1,nrow(multi.df),1)
    multi.df[c(1:3,nrow(multi.df)),]

    
    multi.df$ADA <- as.numeric(multi.df$ADA)
    multi.df$BTC <- as.numeric(multi.df$BTC)
    multi.df$LINK <- as.numeric(multi.df$LINK)
    multi.df$DOT <- as.numeric(multi.df$DOT)
    multi.df$SOL <- as.numeric(multi.df$SOL)
  
    
    multi.df <- na.omit(multi.df)
    
    
    #Calculate Normalized Values for Each Security
#Create an index for each security with values that equal the price of the 	security on each day divided by the security's price on December 31, 	2010.

  
    
    
    multi.df$BTC.idx <- multi.df$BTC/multi.df$BTC[1]
    multi.df$ADA.idx <- multi.df$ADA/multi.df$ADA[1]
    multi.df$LINK.idx <- multi.df$LINK/multi.df$LINK[1]
    multi.df$SOL.idx <- multi.df$SOL/multi.df$SOL[1]
    multi.df$DOT.idx <- multi.df$DOT/multi.df$DOT[1]
    
    options(digits=5)   # this option remains in place until end of session
    multi.df[c(1:3,nrow(multi.df)),]    



```

```{r training & test}

    ADA <- as.data.frame(data.ADA)
    BTC <- as.data.frame(data.BTC)
    LINK <- as.data.frame(data.LINK)    
    DOT <- as.data.frame(data.DOT)    
    SOL <- as.data.frame(data.SOL)
    
    defi <- cbind.data.frame(ADA,BTC, LINK,DOT, SOL)
    summary(defi)
    
    
    defi.train <- defi 
    #Dropping volume variable
    defi.train <- defi.train[-c(6,12,18,24,30)]
    

    ADA$ADA.Open <- as.numeric(unlist(ADA$ADA.Open))
    ADA$ADA.High <- as.numeric(unlist(ADA$ADA.High))
    ADA$ADA.Low <- as.numeric(unlist(ADA$ADA.Low))
    ADA$ADA.Close <- as.numeric(unlist(ADA$ADA.Close))
    ADA$ADA.Adjusted <- as.numeric(unlist(ADA$ADA.Adjusted))  
    
    BTC$BTC.Open <- as.numeric(unlist(BTC$BTC.Open))
    BTC$BTC.High <- as.numeric(unlist(BTC$BTC.High))
    BTC$BTC.Low <- as.numeric(unlist(BTC$BTC.Low))
    BTC$BTC.Close <- as.numeric(unlist(BTC$BTC.Close))
    BTC$BTC.Adjusted <- as.numeric(unlist(BTC$BTC.Adjusted)) 
    
    LINK$LINK.Open <- as.numeric(unlist(LINK$LINK.Open))
    LINK$LINK.High <- as.numeric(unlist(LINK$LINK.High))
    LINK$LINK.Low <- as.numeric(unlist(LINK$LINK.Low))
    LINK$LINK.Close <- as.numeric(unlist(LINK$LINK.Close))
    LINK$LINK.Adjusted <- as.numeric(unlist(LINK$LINK.Adjusted)) 
    
    DOT$DOT.Open <- as.numeric(unlist(DOT$DOT.Open))
    DOT$DOT.High <- as.numeric(unlist(DOT$DOT.High))
    DOT$DOT.Low <- as.numeric(unlist(DOT$DOT.Low))
    DOT$DOT.Close <- as.numeric(unlist(DOT$DOT.Close))
    DOT$DOT.Adjusted <- as.numeric(unlist(DOT$DOT.Adjusted)) 
    
    SOL$SOL.Open <- as.numeric(unlist(SOL$SOL.Open))
    SOL$SOL.High <- as.numeric(unlist(SOL$SOL.High))
    SOL$SOL.Low <- as.numeric(unlist(SOL$SOL.Low))
    SOL$SOL.Close <- as.numeric(unlist(SOL$SOL.Close))
    SOL$SOL.Adjusted <- as.numeric(unlist(SOL$SOL.Adjusted)) 
    
    defi.train$ADA.Open <- as.numeric(unlist(defi.train$ADA.Open))
    defi.train$ADA.High <- as.numeric(unlist(defi.train$ADA.High))
    defi.train$ADA.Low <- as.numeric(unlist(defi.train$ADA.Low))
    defi.train$ADA.Close <- as.numeric(unlist(defi.train$ADA.Close))
    defi.train$ADA.Adjusted <- as.numeric(unlist(defi.train$ADA.Adjusted))
    
    defi.train$BTC.Open <- as.numeric(unlist(defi.train$BTC.Open))
    defi.train$BTC.High <- as.numeric(unlist(defi.train$BTC.High))
    defi.train$BTC.Low <- as.numeric(unlist(defi.train$BTC.Low))
    defi.train$BTC.Close <- as.numeric(unlist(defi.train$BTC.Close))
    defi.train$BTC.Adjusted <- as.numeric(unlist(defi.train$BTC.Adjusted))
    
    defi.train$LINK.Open <- as.numeric(unlist(defi.train$LINK.Open))
    defi.train$LINK.High <- as.numeric(unlist(defi.train$LINK.High))
    defi.train$LINK.Low <- as.numeric(unlist(defi.train$LINK.Low))
    defi.train$LINK.Close <- as.numeric(unlist(defi.train$LINK.Close))
    defi.train$LINK.Adjusted <- as.numeric(unlist(defi.train$LINK.Adjusted))
    
    defi.train$DOT.Open <- as.numeric(unlist(defi.train$DOT.Open))
    defi.train$DOT.High <- as.numeric(unlist(defi.train$DOT.High))
    defi.train$DOT.Low <- as.numeric(unlist(defi.train$DOT.Low))
    defi.train$DOT.Close <- as.numeric(unlist(defi.train$DOT.Close))
    defi.train$DOT.Adjusted <- as.numeric(unlist(defi.train$DOT.Adjusted))
    
     defi.train$SOL.Open <- as.numeric(unlist(defi.train$SOL.Open))
    defi.train$SOL.High <- as.numeric(unlist(defi.train$SOL.High))
    defi.train$SOL.Low <- as.numeric(unlist(defi.train$SOL.Low))
    defi.train$SOL.Close <- as.numeric(unlist(defi.train$SOL.Close))
    defi.train$SOL.Adjusted <- as.numeric(unlist(defi.train$SOL.Adjusted))
    
    
    head(defi.train)
    
    
    defi.training <- defi.train[1:165,]
    defi.test <- defi.training[166:207,]


```



Step 2: Choosing regression method

I am choosing multiple linear regresion for this study based on anecdotal literature studies 
that showed multiple regressions having slightly better accuracy than time-series.
As the price prediction (output) is not categorical, logistic regression may not be the best choice.
Alos, there is no category involved so there is no requirement to create dummy variables

Step 3:Building training and test dataset

As the datasets are in the form of training and test, no need to create a partition.
However, to demonstrate the code to create the partition, code is as follows:

```{r partition}
set.seed(12345)

ADA.tr <- ADA[1:165,]
BTC.tr <- BTC[1:165,]


ADA.ts <- ADA[166:207,]
BTC.ts <- BTC[166:207,]

```

```{r model.ADA}
# building the logistic regression model

regression.model <- lm(ADA.Close  ~ ADA.Open + ADA.High + ADA.Low + ADA.Adjusted,
                data = ADA.tr)

summary(regression.model)
AIC(regression.model)
#[1] -1695.75
BIC(regression.model)
#[1] -1695.75
```
In the summary, we can see that the p-value for Volume > 0.05 which is why we can exclude that variable from
the model. 

Step 6. Predictions
Make predictions

```{r predict}

lm.pred <- predict(regression.model, ADA.ts, type = "response")

lm.pred
summary(lm.pred)


```

Step 7. Evaluation
MSE = Mean Square Error MAE = Mean Absolute Error RMSE = Root Mean Square Error R2 = R-square

```{r Eval}

data.frame(
  R2 = rsquare(regression.model, data = ADA.ts),
  RMSE = rmse(regression.model, data = ADA.ts),
  MAE = mae(regression.model, data = ADA.ts),
  MSE = mse(regression.model, data = ADA.ts)
)

#outputs:
# R2 = 0.9989971
# 
# RMSE =  0.1352867
# 
# MAE = 0.05647035
# 
# MSE = 0.01830249

#create scatterplot of data
plot(ADA.ts$ADA.Close)

plot(regression.model)

```

The MLR model has an r-squared value of 0.9989971, an RMSE of 0.1352867, an MAE of 0.05647035 and an MSE of 0.01830249


Step 8. Results

The multiple regression model (MLR) was built on the variables open, low and High for predicting the closing price for Decentralized finance (ChainLink). 
The regression model has an r2 value of 0.9989971.
The equation to preidict price is:
Price(next day) = OPen(-6.962e-01) + High(8.961e-01) + Low(7.944e-01).
or Price(next day) = -0.6962(Open) + .8961(High) + -.7944 (Low)

As per the visualization, we can see that the model fits the data for most of the predictions. 
The Q-Q plot is seen to be normal as most of the observations follow the reference line.
The residual v fitted plot also shows that the predicted values and observed values are a close fit. 


Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept) -1.160e-03  4.144e-03  -0.280    0.780    
Open        -6.962e-01  1.828e-02 -38.075   <2e-16 ***
High         8.961e-01  1.761e-02  50.895   <2e-16 ***
Low          7.944e-01  1.718e-02  46.244   <2e-16 ***
