---
title             : "Predicting volatility and price of Decentralizezd Finance (DeFI)"
shorttitle        : "Predicting volatility and price of Decentralizezd Finance (DeFI)"

author: 
  - name          : "Gaurav Gade"
    affiliation   : "1"
    
abstract: |
  This study reveals a detailed comparison between several statistical and advanced data science model in
  predicting the price of Decentralized finance toke in addition to Bitcoin. The predictive power of the models
  is tested using both statistical model and advanced machine learning models while considering the time series,
  seasonal trends, based on closing prices, market cap and volume. ML techniques were optimized with the
  addition of variables, layers and using other sophisticated algorithms like Longterm Short term memory.

keywords          : "Defi, Cryptocurrency, Price prediction, Time-series,LSTM"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences, include = FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```
Decentralized Finance is a relatively new concept in crypto where users can borrow or lend their tokens in order
to earn interest or open new position. A majority of research studies and experiments have been conducted on top
10 cryptocurrencies like Bitcoin and Ethereum. However, there has not been a conclusive study of price 
prediction for Decentralized Finance tokens which will be the main focus of my research. The techniques and 
methods that researchers have used to analyze Bitcoin and Ethereum can be expanded to DeFi tokens which is why 
it is important to perform an in-depth study the results of these models. Accordingly,  have I performed a 
literature survey of over 30 articles published in established journals to prepare the foundation of my research
study. 

#   Literature Review

###   Review of traditional statistical methods

 Bayesian Regression and GLM/Random forests can be used to predict the price
of Bitcoin. [@velankar2018bitcoin]. To run the model, the data need to be segment into consecutive intervals of 
sizes 180s, 360s and 720s. Then k-means clustering can be applied to determine the best clusters. Using Bayesian
regression, the second set of prices to calculate the corresponding weights of features can be calculated. Price
change at a specific time interval for time-series data can be used by running GLM/Random forest on two separate
time series data to get two linear models which can then be used to predict the price change. The GARCH-MIDAS 
model allows us to link the daily observations on stock returns with macroeconomic variables, sampled at lower 
frequencies, to determine the impact on the stock volatility. [@hou2013importance]. Another study used this 
model for predicting the volatility the crypto market. According to [@fang2020economic], the study provides new 
evidence about the impact of the volatility in crypto. This paper investigates how the price of cryptocurrency 
is impacted by NVIX or News based implied volatility. Using the GARCH-MIDAS model, it was noted that NVIX has a 
significant negative effect on the price. This leads us to believe that investors in Cryptocurrency get affected
by negative news about crypto. In the paper, [@ma2020cryptocurrency], authors have predicted the accuracy of the
realized variance of Bitcoin. They concluded that the novel MRS-MIDAS model showed an improved accuracy in the 
forecasting of the RV of Bitcoin for 2-week and 1 month time intervals. In this paper, [@conrad2018long] have 
used Long term and Short term volatility to predict Bitcoin volatility. Using the GARCH-MIDAS model, they S&P 
500 companies realized volatility to have a negative and highly significant effect on Bitcoin. S&P 500 risk 
volatility had positive effect on Bit coin’s volatility that can be used to improve long-term forecasts. Per 
[@chen2020bitcoin]), it is possible to predict Bitcoin price using traditional statistical methods like 
Regressions and LDA. However, the study included ML algorithms and based on the results, the researchers
concluded that highly specialized ML techniques were better at performing price prediction. Lastly, the paper - 
[@caporale2019day] studied the effect of certain days in cryptocurrency using traditional statistical tests 
including Student’s t-test, ANOVA, Kruskal-Wallis and regression. The result determined that for Bitcoin, 
Mondays are statistically likely to provide higher returns than other days of the week. No effect was present 
for other cryptocurrency

###   Review of sophisticated algorithms like ML and Neural networks

  According to [@chen2020bitcoin]), we can analyze crypto stocks using ML methods like random forests, XGBoost, 
Quadratic LDA, and SVM by using high-dimension features including property, network, trading, market, attention 
and gold spot price.
The Logistic Regression and Linear Discriminant Analysis for Bitcoin daily price prediction with
high-dimensional features achieved an accuracy of 66 while achieved 67.2% accuracy for the ML methods.
Based on the findings, statistical methods proved to be an easy way to analyze time-series data. However, there 
are other algorithms that outperformed these traditional methods. Since, the data used to analyze cryptocurrency
is in the form of time-series, I researched articles that used the time-series model to predict prices. 
[@kumar2020predicting] has predicted the trends of price for Ethereum based on deep learning techniques 
particularly time-series. 
ARIMA model was found to be one of the easiest and effective machine learning algorithms where time-series data 
is involved. [@ji2019comparative] However, ARIMA model for BTC price prediction resulted in large MSE values. 
It can be used for price prediction in sub-periods of the timespan by dividing the timespan over smaller 
subsets. Deep learning techniques such as multi-layer perceptron (MLP) & long short-term memory (LSTM) can 
help in predicting the price trends of Ethereum. Before building complex models, it is worth-while to check 
whether the inbuilt models available in Python and R libraries could be used for the prediction. Scikitlearn and
Keras libraries have powerful in-built methods for price prediction of Bitcoin at 1 minute intervals. In 
Sci-kit, the ‘Theil-Sen Regression’ method and ‘Huber Regression’ method had a MSE of 0.000375 and 0.000373 
respectively while the R2 was 99.2%.  [@phaladisailoed2018machine]. For deep learning based regression models, 
Keras library can be used to create LSTM and GRU models with MSE  0.000431 and 0.00002 respectively. The R2 was 
99.2% in both cases.  [@phaladisailoed2018machine]. Hence, we can utilize these methods to effectively predict 
prices. Several research papers have used neural networks for prediction. Per [@sin2017bitcoin] ANN can be used 
to predict next day price movement (positive or negative) of Bitcoin with an accuracy of 64%. The paper explores
the relationship between the features of Bitcoin and the next day change in the price of Bitcoin using an 
Artificial Neural Network ensemble approach called Genetic Algorithm based Selective Neural Network Ensemble 
using 5 Multi-Layered Perceptron (MLP). [@azari2019bitcoin] has used time-series data to develop neural networks
like recurrent neural networks, convolutional neural networks, and autoregressive integrated moving average 
(ARIMA) to develop a predictive model for Bitcoin price. RNN’s were found to significantly outperform ARIMA in 
predictive accuracy. However, according to a majority of articles, it seems that the LSTM models from machine 
learning were most effective in price prediction. Per experimental results conducted by [@ji2019comparative] 
LSTM-based prediction models slightly outperformed the other prediction models for regression problems, 
DNN-based prediction models performed the best for classification problems. Per [@dutta2020gated], Sequence 
model can be applied with a ﬁxed set of exogenous and endogenous factors for forecasting future crypto prices. 
The ML models like RNN and LSTM have been known to preform better than traditional time-series models. 
[@dutta2020gated]. [@patel2020deep] Utilizes LSTM (Long term Short term memory) a, Gated recurrent unit and 
Neural networks to predict crypto prices specifically Lite coin and Monero. The algorithm was able to predict 
with a high degree of accuracy the prices of these two currencies. Per [@pant2018recurrent] tweets related to 
Bitcoin can be fed to RNN model along with historical price to predict the price for next time frame. The 
accuracy for sentiment classification of tweets in two class positive and negative is found to be 81.39 % and 
the overall price prediction accuracy using RNN is found to be 77.62%.  Per [@jay2020stochastic] we can utilize 
a neural network model for the prediction of cryptocurrency. By using a layer based model for the observed 
feature activation of neural network to simulate the market volatility. The model used was MLP (multi-layer 
perceptron) and Long-term short memory. These models showed an improved accuracy than deterministic models to 
predict the price of 3 major currencies like Bitcoin, Ethereum and Lite coin. In [@sebastiao2021forecasting], 
researchers were able to develop strong and flexible techniques to discover predictability of major 
cryptocurrencies and formulate profitable trading strategies by using samples during the ‘bear’ market. The 
trading strategies were developed using the machine learning models and were validated and tested during the 
market fluctuations allowing the study to have an inference of its true predictions between the validation and 
test period. The success rates of individual machine learning models and the conclusive positive results with an
achievable performance of major cryptocurrencies using model assembling have helped support the study.

###   Review of sentiment analysis of twitter and other social media platforms
  
  According to the research conducted in [@abraham2018cryptocurrency], Google Trends and tweet 
volume were found to be highly correlated with price. When prices are falling twitter sentiment is not be an 
effective indicator.  Based on the research in [@stenqvist2017predicting], VADER (Valence Aware Dictionary and 
Sentiment Reasoner) using Random Forest regression can be used to predict twitter sentiment and to identify 
relationships between different kinds of input. Twitter sentiments can be used in real-time for crypto price 
prediction. Using a Spark based architecture that can handle large volume of data, sentiment analysis can be 
performed in real-time to build and utilize a predictive model that can adjust the weights to accurately provide
timely insights that can be used to make a decision. [@mohapatra2019kryptooracle]. Another paper - 
[@misnik2018neural] estimates the price of Bitcoin using market data by analyzing social and time factors. 
Multilayer perceptron was used along with LSTM neural networks. For time-series, LSTM was considered to be the 
best approach. (Misnik, (2018)).  News and Social media can be used to effectively predict the price 
fluctuations of Bitcoin, Lite coin and Ethereum. According to [@lamon2017cryptocurrency], the study utilizes 
traditional supervised learning algorithms for text based classification. Daily news and social media data was 
labelled based on price fluctuations. That way price fluctuation can be predicted without having to first 
predict the sentiment. The model can be used to predict the largest % increase or decrease for Bitcoin and 
Ethereum.  Per [@wolk2020advanced] tweet frequency had a high inverse correlation with crypto prices. Bad news 
especially caused an increase in post/tweet frequency. The study looks at price prediction for crypto currency 
using SVM, SGD, GBM, MLP Neural network, Least squares linear regression, ADAboost and Bayesian Ridge 
Regression, Decision Tree and ElasticNet. A hybrid model built on the mean of all the models was used to make 
the final prediction. In this article [@nizzoli2020charting] have discussed about the manipulation of 
cryptocurrencies by influencers like Twitter, Telegram and Discord and how topic modeling techniques were used 
to detect the fraud schemes. Major data set of messages was used as samples to find the sham schemes categorized
in 2 sets ‘pump and dump’ and ‘Ponzi’ majorly inclining towards Telegram accounting to 20% of the total 
channels. In addition to this, out of total shared Twitter invite links for bot detection 93% were pointing to 
Telegram channels. The researcher predicts to fight this manipulation and cryptocurrency price abuse by using 
such records. [@kraaijeveld2020predictive] used sentiment analysis approach to predict the price returns of 
largest cryptocurrencies. He has discussed about Twitter sentiment being used, along with financial data and 
granger-causality testing to predict the bullishness ratio and thus find the predictive power for some 
cryptocurrencies. Moreover this approach also led to findings of at least 1-14% of obtained tweets tweeted by 
‘bot’ accounts on Twitter.

###   Review of Volatility of price and fluctuation
  
  Several studies focused on predicting the volatility rather than price movement. The reason is that volatility
  based on the direction of price movement could result in huge gains or losses. In order to optimize the 
portfolio for risk mitigation, it is important to understand how volatile the market it. In this research paper,
[@catania2018predicting] have studied the conditional volatility for Bitcoin, Ethereum, Lite coin and Ripple. 
They determined that the volatility in the crypto market was found to be similar in price movement to other 
financial time-series like foreign exchange returns. They were able to determine the effect of accounting for 
long-memory in the volatility process.  In another study by [@yi2018volatility], researchers attempted to study 
the correlation of cryptocurrencies using static and dynamic volatility connectedness. The technique used was 
LASSO-VAR for estimation. 52 cryptocurrencies were found to be tightly interconnected. Finally, in the article, 
[@walther2019exogenous], the researchers have used a mixed data sampling approach to forecast the volatility of 
highly capitalized cryptocurrencies and its index. They have discussed the external factors driving this 
volatility major being Global Real Economic Activity factor which has been proved effective for the market 
fluctuations. Moreover, the derived low loss functions due to average forecasting shows us that the factors are 
time varying and that the model averaging approach broadens the impact of single driving factors.

###   Review of DEFI - decentralized finance 
  
  DEFI is a new way of borrowing or lending money in crypto without having to go to traditional banking methods 
like banks, brokerages or exchanges. DeFI can also provide good returns on investment. However, the DEFI space 
is still in its infancy and susceptible to risks. In the article, [@gudgeon2020decentralized], researchers 
identify two weaknesses in DeFi protocols that are susceptible to price abuse in Crypto. First being over 
collateralization, author discusses how DeFi lending protocols’ flexibility be measured during the fall of its 
assets and show the speed at which DeFi protocol would become under Collateralized. Second being a governance 
attack, the author discusses how attacker is able to steal all the collateral. In this paper, [@corbet2021defi] 
Found out strong correlation between Bitcoin and DeFi tokens. Ethereum or Bitcoin do not cause bubbles in the 
DeFi market according to their research. Mainly, the cryptocurrency Chain link and Maker are the main drivers. 
Therefore, while perform price predictions on DeFi, they suggest treating DeFi tokens as separate asset classes.

#   Methods

  In this study, the main topic is to analyze DeFi cryptocurrency. I started with statistical 
modeling tests on the top 4 Decentralized Finance tokens - ADA, SOL, LINK and DOT in addition to Bitcoin. Based 
on the literature survey, it is evident that neural networks and Machine Learning algorithms like long-term 
short-term memory models outperform other techniques to analyze crypto market. In addition to build these 
models, I used time-series modeling and built random forest for the the purpose of forecasting. 
I used in-built functions and libraries from R as well as some core libraries like Keras and Tensorflow as these
have proved to be extremely effective in price prediction and 
easy to build and manage. [@phaladisailoed2018machine]. In addition to the time-series data, LSTM neural 
networks was implemented using recurrent neural network algorithm. For data preparation, the historic datasets 
available from sources like Yahoo finance or Coinmarketcap were converted to time-series data for processing. 
Iformation without proper actionable insights is not that effective. Therefore, I will be using data 
visualizations to showcase DEFI prices and predicted values. 
I have performed a comparative evaluation of the result 

#    Participants
  The main requirement for my analysis is crypto stock data and twitter data scraped from the internet. Based on
  the time-series data available, I can run my predictive models. In addition, the neural networks like 
convolutional neural network, recurrent neural network are memory intensive programs that may require processing
power of a higher magnitude to run in a reasonable amount of time. I am therefore planning to use resources like
cloud (Microsoft Azure or Amazon AWS). This will ensure that I can scale my program and deal with large 
datasets. 

#   Procedures
  The study would start with statistical modeling tests on Defi tokens like DOT, ADA, SOL and LINK. Since 
this is time-series data, I ran time forecasting methods. Then, I have built  a LSTM neural network to perform a
basic crypto price  prediction. I used in-built functions and libraries from Keras and Sci-kit in Python. 
Then I would be plotting and developing proper visualization of the Defi prices. 

#   Material
  To build this solution, I will be utilizing Keras, Sci-kit and Tensorflow libraries,
  Long-Term Short memory and Neural networks like CNN, RNN.  ARIMA, Random Forest and other time-series 
  forecasting models were used.

#   Measures

  For time-series modeling, the study will use measures like Autocorrelation values, partial autocorrelation, 
and forecasting accuracy using measures like MAPE to determine effectiveness.
For LSTM and machine learning algorithms, I will be using a split dataset for validation. Accuracy of the model 
will be used to determine the best performing model. 

#   Data analysis

  For running the time-series, the following analysis steps were performed: 
  1.    Plotting the autocorrelation to see if there are many lags in the time series. 
  2.    Creating validation test-datasets. Using time-intervals to capture maximum variance and volatility. 
  3.    Building ARIMA and SARIMA model for price prediction. 
  4.    Running Random Forests

  For LSTM, I performed following steps:
  1.    Create training and test values, scaling data and assigning training values	
	2.    Create the model with appropriate parameters like optimizers, epochs
	3.    Fit the model to the training data
	4.    Assign test and predicted values and then plotting the results

  For PCA, I have perfomed following steps:
  1.    Building the covariance matrix using pearson co-efficent on the data
  2.    Using the princomp function to build the PCA model
  
  For Cluster Analysis, I have performed the following steps:
  1.    Find optimal number of clusters using elbow method 
  1.    calculate the average silhouette distances for 2 to n-1 clusters 
  3.    Perform the K-Means cluster analysis
  4.    Visualize cluster
  5.    Build hierarchical cluster



#   Results

#   Analysis

##    Exploratory Data Analysis

`render(input = "C:/Users/nagar/Documents/ANLY 699 final project/ANLY 699 Time Series [26 Nov 2021].rmd", output_format = "pdf_document")`

1. Import data for DeFi Tokens and BTC - BTC, CARDANO, CHAINLINK, POLKADOT, SOLANA

    

I used `r cite_r("r-references.bib")` for all the analysis.


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
